---
title: "MA 331 homework 1"
author: "Josh Gribbon"
output:
  html_document: default
  pdf_document: default
---
# Data
(0.2, 1.1), (1.2, 2.3), (0.9, 1.1), (2.2, 3.6), (6.2, 0.1), (3.6, 4.8), (5.3, 6.5), (6.2, 7.8), (7.1,8.0), (8.6, 9.4), (9.0,9.8)
```{r}
x <- c(0.2, 1.2, 0.9, 2.2, 6.2, 3.6, 5.3, 6.2, 7.1, 8.6, 9.0)
y <- c(1.1, 2.3, 1.1, 3.6, 0.1, 4.8, 6.5, 7.8, 9.0, 9.4, 9.8)
df <- data.frame(x, y)
```

# 1 - Plot histogram, pie chart and describe the distribution of x and y.
```{r}
hist(x,main="Histogram of X")
hist(y,main="Histogram of Y")
pie(x,main="Pie chart of X")
pie(y,main="Pie chart of Y")
```

Data set X is not symmetric, and is fairly spread out, with values between 0 and 10.

Data set Y is symmetric and very spread out, with values between 0 and 10.


# 2 - For x and y, build the box-plot, compute their five-number summaries and variances. Are there any outliers of x and y respectively?
```{r}
boxplot(x, main="Boxplot of X")
boxplot(y, main="Boxplot of Y")
```
```{r}
summary(x)
summary(y)
```
```{r}
var(x)
var(y)
```

Neither X or Y have any notable outliers.

# 3 - Obtain the scatter plot of (x,y) and evaluate their correlation coefficient. Describe the linear association between x and y.
```{r}
plot(x, y, main="Scatterplot of X vs Y")
cor(x, y)
```

X and Y have a positive linear correlation, and are fairly strongly correlated.

# 4 -  Are there any outliers of (x,y)? If yes, remove them and compute the correlation coefficient again. What difference do you observe?
Yes, (6.2, 0.1) is an outlier.
```{r}
x2 <- c(0.2, 1.2, 0.9, 2.2, 3.6, 5.3, 6.2, 7.1, 8.6, 9.0)
y2 <- c(1.1, 2.3, 1.1, 3.6, 4.8, 6.5, 7.8, 9.0, 9.4, 9.8)
plot(x2, y2, main="Scatterplot of X vs Y with outliers removed")
cor(x2, y2)
```

The correlation increased after removing the outliers.
